---
title: "Figures for paper"
author: "JNielsen"
date: "13/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction

This Rmarkdown produces the summary statistics and graphical outputs presented in the manuscript "Evaluation of experimental considerations in using acute heat stress assays to quantify coral thermal tolerance" by Nielsen et al 2022. This document follows the layout from the manuscript. 

Libraries required for these analyses
```{r Libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(effects)    # for partial effects plots
library(ggeffects)  # for effects plots in ggplot
library(patchwork)  # for arranging multiple plots
library(ggplot2)    # for graphing
library(gridExtra)
library(ggpubr)     # for using ggarrange() function to organise plots
library(extrafont)  # for setting fonts in ggplot
loadfonts(device = "win") # needed to apply the fonts
library(jpeg) # for including images in patchwork figures
library(ggpattern) # for drawing patterns on the box plots
library(lattice) # for the correlation matrics
library(reshape2) # for the metling of the correlation matrix

# for maps
#install.packages("remotes")
#remotes::install_github("https://github.com/open-AIMS/gisaimsr")
library(raster) # for map
library(ggspatial) 
library(sf)
library(dataaimsr)
library(gisaimsr)
library(ggrepel)
library(rgeos)
library(rnaturalearth)

# for RDA 
library(vegan)
library(GGally)
library(corrplot)
library(mvabund)
library(scales) # for viridis palette
```

# Map
Map of the 7 reefs included in this study. 

```{r}
coordsmap <- read.csv("coordsmap.csv")

cook <- data.frame(cooktown = ("Cooktown"), lat = -15.47743, lon = 145.24924)

FNGBR = ggplot() +
  geom_sf(data = gbr_feat, lwd = 0.01, fill = "grey90") +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true",
      pad_x = unit(0.25, "in"), pad_y = unit(0.4, "in"),
      style = north_arrow_fancy_orienteering) +
  coord_sf(xlim = c(142, 146.75), ylim = c(-16, -10.5)) +
  geom_point(data = coordsmap, mapping = aes(x = lon, y = lat, size = n, colour = MMM)) +
  geom_label_repel(data = coordsmap, 
                   mapping = aes(x = lon, y = lat,
                                 label = reef),
                   hjust = 0, vjust = 0,
                   xlim = 146, 
                   direction = "both",
                   min.segment.length = 0.5,
                   segment.size = 0.2,
                   size = 3) +
  geom_text_repel(data = cook, 
                   mapping = aes(x = lon, y = lat,
                                 label = cooktown),
                   hjust = 0, vjust = 0) +
  theme_classic() +
    theme(panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank()) +
  labs(x = "Longitude",
       y = "Latitude",
       title = "",
       subtitle = "") +
  scale_color_distiller() #palette = "Spectral"
FNGBR

oz <- ne_countries(country = "Australia", scale='medium',returnclass = 'sf')

inset = ggplot() +
  geom_sf(data = oz, fill = "grey30", col = "transparent") +
  geom_sf(data = gbr_bounds, lwd = 0.1, col = "grey30", fill = "transparent") +
  theme_void() +
  theme(legend.position = "none") 

annotated_map = FNGBR + 
  annotation_custom(ggplotGrob(inset), xmin = 143.4, xmax = 144.5, ymin = -16, ymax = -14.5)
annotated_map

#ggsave("HeatStressMethodsMap.jpeg")
```

# Summary statistics for size effects

Getting means, standard error for all variables in size section 
First execute the summarySE function. 
```{r}
## Summary Function (SE) - execute this entire block of code to store.
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  require(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary; it's not easy to understand...
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun= function(xx, col, na.rm) {
                   c( N    = length2(xx[,col], na.rm=na.rm),
                      mean = mean   (xx[,col], na.rm=na.rm),
                      sd   = sd     (xx[,col], na.rm=na.rm)
                   )
                 },
                 measurevar,
                 na.rm
  )
  
  # Rename the "mean" column    
  datac <- rename(datac, c("mean"=measurevar))
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}
```

```{r}
# Colour change
all.summ.Col = summarySE(Size,measurevar="ColourChange",groupvars=c("Species","Treatment", "Size"))
all.summ.Col

# Chlorophyll 
datCleanChla = Size[complete.cases(Size$Chla_ug_cm2),] # clean the variables first if there are any NAs
all.summ.Chla = summarySE(datCleanChla,measurevar="Chla_ug_cm2",groupvars=c("Species","Treatment", "Size"))
all.summ.Chla

# Catalase 
datCleanCat = Size[complete.cases(Size$CatalaseU),] # clean the variables first if there are any NAs
all.summ.Cat = summarySE(datCleanCat,measurevar="CatalaseU",groupvars=c("Species", "Treatment","Size"))
all.summ.Cat

# Protein
datCleanProt = Size[complete.cases(Size$Protein_ug_cm2),] # clean the variables first if there are any NAs
all.summ.Prot = summarySE(datCleanProt,measurevar="Protein_ug_cm2",groupvars=c("Species","Treatment", "Size"))
all.summ.Prot

all.summ.ProtSize = summarySE(datCleanProt,measurevar="Protein_ug_cm2",groupvars=c("Species", "Size"))
all.summ.ProtSize

all.summ.ProtSizeNoSpecies = summarySE(datCleanProt,measurevar="Protein_ug_cm2",groupvars=c("Size"))
all.summ.ProtSizeNoSpecies


# Fv/Fm
datCleanPhot = SizePlotFvFm[complete.cases(SizePlotFvFm$value),] # clean the variables first if there are any NAs
all.summ.Phot = summarySE(datCleanPhot,measurevar="value",groupvars=c("species","treatment", "size"))
all.summ.Phot

```

## Mean surface area

Calculate mean surface areas for both sizes and species, and split by species
```{r}
datCleanSurface = Size[complete.cases(Size$Surface),] # clean the variables first if there are any NAs
all.summ.Surface = summarySE(datCleanSurface,measurevar="Surface",groupvars=c("Species", "Size"))
all.summ.Surface

AtenSize <- Size %>% 
  filter(Species == "A. tenuis")

PdamSize <- Size %>% 
  filter(Species == "P. damicornis")
```


# Figure 1 Composite

Needs to incorporate the map from above, time lines for size and normal experiments, plus the "time" time line

Experimental profile, normal and size experiments
```{r}
normaldesign <- read.csv("Fig1NormalTimeFramedesign.csv")
head(normaldesign)

normaldesign$Treatment <- factor(normaldesign$Treatment, levels = c('Ambient', 'Mid', 'High'))
```
Graph the normal experiment
```{r}
g4 <- ggplot(normaldesign, aes(x = Time, y = Temperature, colour = Treatment))+
  geom_line(size = 1)+
  geom_segment(aes(x = 18, y = 30, xend = 18, yend = 29), 
               color = "black", size = 0.5, ends = "none")+
  annotate(geom = "text", x = 17.6, y = 30.05, label = "Sampling", 
           hjust = 0, vjust = 0.5, angle = 70, size = 3)+
  geom_segment(aes(x = 17, y = 30, xend = 17, yend = 29), 
               color = "grey40", size = 0.5, ends = "last")+
  annotate(geom = "text", x = 16.9, y = 30.05, label = "PAM", 
           hjust = 0, vjust = 0.5, angle = 70, size = 3, colour = "grey40") +
  scale_colour_manual(values = c("blue2", "yellow3", "red"))+
    ylab(bquote('Temperature (°C)'))+
    xlab(bquote('Time (h)'))+
    theme_classic()+
  theme(strip.text = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "right")
```
Graphing the time sampling design
```{r}
timedesign <- read.csv("Fig1TimeSeries.csv")
timedesign$Treatment <- factor(timedesign$Treatment, levels = c('Ambient', 'High'))

g5 <- ggplot(timedesign, aes(x = Time, y = Temperature, colour = Treatment))+
  geom_line(size = 1)+
  geom_segment(aes(x = 7.8, y = 30, xend = 7.8, yend = 29), #T0
               color = "black", size = 0.5, ends = "last")+
    annotate(geom = "text", x = 7.8, y = 30.05, label = "T0", 
           hjust = 0, vjust = 0.5, angle = 70, size = 3)+
  geom_segment(aes(x = 9.5, y = 30, xend = 9.5, yend = 29), #T1
               color = "black", size = 0.5, ends = "last")+
    annotate(geom = "text", x = 9.5, y = 30.05, label = "T1", 
           hjust = 0, vjust = 0.5, angle = 70, size = 3)+
  geom_segment(aes(x = 13.5, y = 30, xend = 13.5, yend = 29), #T2
               color = "black", size = 0.5, ends = "last")+
    annotate(geom = "text", x = 13.5, y = 30.05, label = "T2", 
           hjust = 0, vjust = 0.5, angle = 70, size = 3)+
  geom_segment(aes(x = 17.5, y = 30, xend = 17.5, yend = 29), #T3
               color = "black", size = 0.5, ends = "last")+
    annotate(geom = "text", x = 17.5, y = 30.05, label = "T3", 
           hjust = 0, vjust = 0.5, angle = 70, size = 3)+
  geom_segment(aes(x = 21.5, y = 30, xend = 21.5, yend = 29), #T4
               color = "black", size = 0.5, ends = "last")+
    annotate(geom = "text", x = 21.5, y = 30.05, label = "T4", 
           hjust = 0, vjust = 0.5, angle = 70, size = 3)+
  geom_segment(aes(x = 31.5, y = 30, xend = 31.5, yend = 29), #T5
               color = "black", size = 0.5, ends = "last")+
    annotate(geom = "text", x = 31.5, y = 30.05, label = "T5", 
           hjust = 0, vjust = 0.5, angle = 70, size = 3)+
  geom_segment(aes(x = 55.5, y = 30, xend = 55.5, yend = 29), #T6
               color = "black", size = 0.5, ends = "last")+
    annotate(geom = "text", x = 55.5, y = 30.05, label = "T6", 
           hjust = 0, vjust = 0.5, angle = 70, size = 3)+
  scale_colour_manual(values = c("blue2", "red"))+
    ylab(bquote('Temperature (°C)'))+
    xlab(bquote('Time (h)'))+
    theme_classic()+
  theme(strip.text = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "right")
```


Arrange the plots altogether
```{r}
ggarrange(g1,
          ggarrange(g4, g5, nrow = 2, common.legend = TRUE, legend = "bottom"), ncol = 2)

right.panel = ggarrange(g4, g5, nrow=2, common.legend =TRUE, legend = "bottom",
                        labels = c ("B", "C"), 
                        font.label = list(size = 10, 
                                          colour = "black", face = "bold", 
                                          family = "Helvetica", position = "top"))

g6 <- ggarrange(g1, right.panel, ncol = 2, widths = c(1.4,1),
          labels = c("A"), 
                        font.label = list(size = 10, 
                                          colour = "black", face = "bold", 
                                          family = "Helvetica", position = "top"))

g6
ggsave("Fig1Comp.jpg", g6, width = 25, height = 15, units = "cm")
```



# Size plots 

First load up the datasets
# Size analysis 
```{r}
# Load size data for both species, set factors
Size <- read.csv("Size.csv") %>%
  mutate(SampleID = as.factor(SampleID),
         Species = as.factor(Species),
         ClipID = as.factor(ClipID),
         RackID = as.factor(RackID),
         GenotypeID = as.factor(GenotypeID),
         Treatment = as.factor(Treatment),
         Tank = as.factor(Tank),
         Size = as.factor(Size))

Size$TreatSize <- paste(Size$Size, Size$Treatment) # for significant interaction terms later
Size$TreatSize = as.factor(Size$TreatSize)
summary(Size)
head(Size)
```

Preparing data for plotting
```{r}
SizePlot <- Size %>% 
  dplyr::select(-c(BeforeColour, AfterColour, Surface, TreatSize, Zoox_ug_protein)) %>% # unselecting certain columns I don't want to plot.
  gather(key = variable, value = value, -c(SampleID, GenotypeID, Species, ClipID, RackID, Treatment, Tank, Size)) 
head(SizePlot) # sanity check of new data frame
```

Colour information for the viridis palette used to keep graphs consistent
```{r}
show_col(viridis_pal()(5))
# purple = #440154FF = colour change
# blue = #3B528BFF = chlorophyll
# Teal = #21908CFF = FvFm
# Green = #5DC863FF = Protein
# Yellow = #FDE725FF = Catalase

```


```{r}
# Catalase, both species

SizePlotCat <- Size %>% 
  dplyr::select(-c(BeforeColour, AfterColour, Surface, TreatSize, Zoox_ug_protein)) %>%
    filter(CatalaseU < 6) %>%
  gather(key = variable, value = value, -c(SampleID, GenotypeID, Species, ClipID, RackID, Treatment, Tank, Size)) 
head(SizePlotCat)


g5 <- ggplot(SizePlotCat %>% filter(variable=="CatalaseU"), 
             aes(x = Treatment, y = value, fill = Size, pattern = Size)) +
  geom_boxplot()+
  geom_boxplot_pattern(position = position_dodge(preserve = "single"),
                       colour = "black",
                       pattern_fill = "black",
                       pattern_angle = 45,
                       pattern_density = 0.25,
                       pattern_spacing = 0.025,
                       pattern_key_scale_factor = 0.6)+
  scale_pattern_manual(values = c(Small = "stripe", Large = "none"))+
  scale_fill_manual(values = c("#440154FF", "#440154FF"))+
  facet_wrap(variable ~ Species, scales = "fixed", ncol = 2) +
  labs(y = "Catalase activity U", x = "Treatment") +
  theme_classic()+
  theme(axis.title.y = element_text(family = "serif", size = 8),
        axis.title.x = element_text(family = "serif", size = 10),
        strip.text = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "none")

```


```{r}
# Chlorophyll, both species

g3 <- ggplot(SizePlot %>% filter(variable=="Chla_ug_cm2"), 
             aes(x = Treatment, y = value, fill = Size, pattern = Size)) +
  geom_boxplot()+
  geom_boxplot_pattern(position = position_dodge(preserve = "single"),
                       colour = "black",
                       pattern_fill = "black",
                       pattern_angle = 45,
                       pattern_density = 0.25,
                       pattern_spacing = 0.025,
                       pattern_key_scale_factor = 0.6)+
  scale_pattern_manual(values = c(Small = "stripe", Large = "none"))+
  scale_fill_manual(values = c("#3B528BFF", "#3B528BFF"))+
  facet_wrap(variable ~ Species, scales = "fixed", ncol = 2) +
  ylab(bquote('Chl-a (' *mu~g ~cm^-2*')'))+
  theme_classic()+
  theme(axis.title.y =  element_text(family = "serif", size = 8), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "none")
```


```{r}
# Colour change

g1 <- ggplot(SizePlot %>% filter(variable=="ColourChange"), 
             aes(x = Treatment, y = value, fill = Size, pattern = Size)) +
  geom_boxplot()+
  scale_fill_manual(values = c("#FDE725FF", "#FDE725FF"))+
  geom_boxplot_pattern(position = position_dodge(preserve = "single"),
                       colour = "black",
                       pattern_fill = "black",
                       pattern_angle = 45,
                       pattern_density = 0.25,
                       pattern_spacing = 0.025,
                       pattern_key_scale_factor = 0.6)+
  scale_pattern_manual(values = c(Small = "stripe", Large = "none"))+
  facet_wrap(variable ~ Species, scales = "fixed", ncol = 2) +
  labs(y = "Colour change", x = "")+
  theme_classic()+
 theme(axis.title.y =  element_text(family = "serif", size = 8), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "none")
```


```{r}
# Protein
SizePlotProt <- Size %>% 
  dplyr::select(-c(BeforeColour, AfterColour, Surface, TreatSize, Zoox_ug_protein)) %>% 
    filter(Protein_ug_cm2 > -0.000000000000000001) %>%
  gather(key = variable, value = value, -c(SampleID, GenotypeID, Species, ClipID, RackID, Treatment, Tank, Size)) 
head(SizePlotProt)


g4 <- ggplot(SizePlotProt %>% filter(variable=="Protein_ug_cm2"), 
             aes(x = Treatment, y = value, fill = Size, pattern = Size)) +
  geom_boxplot()+
  scale_fill_manual(values = c("#5DC863FF", "#5DC863FF"))+
  geom_boxplot_pattern(position = position_dodge(preserve = "single"),
                       colour = "black",
                       pattern_fill = "black",
                       pattern_angle = 45,
                       pattern_density = 0.25,
                       pattern_spacing = 0.025,
                       pattern_key_scale_factor = 0.6)+
  scale_pattern_manual(values = c(Small = "stripe", Large = "none"))+
  facet_wrap(variable ~ Species, scales = "fixed", ncol = 2) +
  ylab(bquote('Protein content ('*mu~g ~cm^-2*')'))+
  xlab("")+
  theme_classic()+
 theme(axis.title.y =  element_text(family = "serif", size = 8), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "none")
```

First prepare PAM data for analysis by calculating Fv/Fm
```{r}
# First prepare PAM data for analysis by calculating Fv/Fm
PAM_Size <- read_csv("PAM_SIZE.csv") %>%
  mutate(genus_species = paste0(substr(genus, start = 1, stop = 1), "_", species),
         tank = as.factor(tank),
         colonyID = as.factor(colonyID),
         replicate = as.factor(replicate),
         species = as.factor(species),
         size = as.factor(size),
         treatment = as.factor(treatment),
         collection.depth = as.numeric(collection.depth),
         Fm = ifelse(Fm <= F, F+1, Fm), # There should not be any Fm values < F
         FvFm = (Fm - F)/Fm)

SizePlotFvFm <- PAM_Size %>% 
  dplyr::select(-c(time_date, genus, site, PAMID, F, Fm, Y, collection.depth, bleaching.score, Region, Sector, Shelf, genus_species, replicate)) %>%
  gather(key = variable, value = value, -c(sampleID, colonyID, species , rack, rack.position, treatment, tank, size)) 
head(SizePlotFvFm)
summary(SizePlotFvFm)


g2 <- ggplot(SizePlotFvFm %>% filter(variable=="FvFm"), aes(x = treatment, y = value, fill = size, pattern = size)) +
  geom_boxplot()+
  scale_fill_manual(values = c("#21908CFF", "#21908CFF"))+
    geom_boxplot_pattern(position = position_dodge(preserve = "single"),
                       colour = "black",
                       pattern_fill = "black",
                       pattern_angle = 45,
                       pattern_density = 0.25,
                       pattern_spacing = 0.025,
                       pattern_key_scale_factor = 0.6)+
  scale_pattern_manual(values = c(Small = "stripe", Large = "none"))+
  facet_wrap(variable ~ species, scales = "fixed", ncol = 2) +
  labs(y = "Photosynthetic efficiency (Fv/Fm)", x = "") +
  theme_classic()+
 theme(axis.title.y =  element_text(family = "serif", size = 8), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "none")
```


```{r warning = FALSE, message = FALSE}
# organising the plots with patchwork

size_plot = g1/g2/g3/g4/g5

ggsave(plot = size_plot, file = "NewNewSizePlot.jpeg", width = 15, height = 20, units = "cm")
```


# Time effect, percent data, point plots with lines fitted

```{r}
TimeSymbiont <- read.csv("TimePlotSymbiont.csv") %>%
  filter(variable != "Zoox") %>%
  filter(ID != "63") # very large outlier in chla

g1 <-ggplot(TimeSymbiont, aes(x = TimeHour, y = value, colour = variable, fill = variable)) +
  geom_point(alpha = 0.5)+
  scale_x_continuous(breaks=c(0,2,6,10,14,24,48),labels = c("0", "2", "6", "10", "14","24", "48"))+
  scale_colour_manual(values = c("#3B528BFF", "#440154FF", "#21908CFF")) +
  scale_fill_manual(values = c("#3B528BFF", "#440154FF", "#21908CFF")) +
  labs(y = " Change Relative to Ambient (%)", x = "Sampling Time Point (h)") +
  geom_smooth(aes(colour = variable), method = "lm")+
  ylim(-130,10)+
  theme_classic()+
  theme(axis.title.y =  element_text(family = "serif", size = 10), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "right", 
        legend.text = element_text(family = "serif", size = 10), 
        legend.title = element_blank())


# holobiont plot
TimePlotHostSeries <- read.csv("TimePlotHostSeries.csv")
summary(TimePlotHostSeries)
TimePlotHostSeries$TimeHour = as.factor(TimePlotHostSeries$TimeHour)

summary(TimePlotHostSeries)
TimePlotHostSeries <- TimePlotHostSeries %>%
  mutate(TimeHour = as.numeric(TimeHour))

g2 <- ggplot(TimePlotHostSeries, 
       aes(x = TimeHour, y = value, colour = variable, fill = variable)) +
  geom_point(alpha = 0.5)+
  scale_x_continuous(breaks=c(0,2,6,10,14,24,48),labels = c("0", "2", "6", "10", "14","24", "48"))+
  scale_colour_manual(values = c("#FDE725FF", "#5DC863FF")) +
  scale_fill_manual(values = c("#FDE725FF", "#5DC863FF")) +
  labs(y = " Change Relative to Ambient (%)", x = "Sampling Time Point (h)") +
  geom_smooth(method = "lm")+
  ylim(-150,10)+
  theme_classic()+
  theme(axis.title.y =  element_text(family = "serif", size = 10), 
        axis.text.x = element_text(family = "serif", size = 10),
        axis.title.x = element_text(family = "serif", size = 10),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "right", 
        legend.text = element_text(family = "serif", size = 10), 
        legend.title = element_blank())

gTime = g1/g2 + plot_layout(guides="collect")

  
#ggsave(plot = gTime, file = "LMNewTimePlot.jpeg", width = 15, height = 15, units = "cm")
```

# Fixing the PCA for publication 

Need to use same five colours as above for the physiology (not sure that can be done)
Legend inside a box
Increase font
Maybe use ambient, mid, high colours from the original sampling design figures


Examining the correlations between traits. 
The larger and more blue dots indicate higher correlation between variables. 
Here, Colour change and FvFm are highly correlated while colour change is also correlated to chl a content and protein content. FvFm is also correlated with chl a and protein content. 
```{r}
tenuis %>% dplyr::select(all_of(physio)) %>% 
  cor %>% 
  corrplot(type = "upper", diag = F)
```

```{r message = FALSE}
ggpairs(tenuis %>% dplyr::select(all_of(physio)), lower = list(continuous = "smooth"), 
        diag = list(continuous = "density"),
        axisLabels = "show")
```


All data code needed for PCA figure. Execute all
```{r}
#Import Data
tenuis1 = read.csv("TenuisMDS.csv") %>% 
  mutate(Reef = as.factor(Reef),
         Species = as.factor(Species),
         Treatment = as.factor(Treatment))

tenuis <- tenuis1 %>% # just getting rid of some long column names
  dplyr::rename(Chla = Chla_ug_cm2, 
                Protein = Protein_ug_cm2) 
 

#outliers
tenuis %>% filter(Reef == "Corbett") %>% arrange(desc(CatalaseU)) # very large catalase outlier in sample 891
tenuis <- tenuis %>% 
  filter(SampleID != "891")

#Isolate traits in a separate data frame 
physio = c("Chla", "Colour.Change", "CatalaseU", "Protein", "FvFm")

# make data frame for RDA
data.rda <- rda(tenuis %>% dplyr::select(all_of(physio)), scale=TRUE)  # based on correlation, pca is an unconstrained redundancy analysis. 
summary(data.rda, display=NULL)

# store data, not sure what difference between sites and species is.
data.sites.scores <- data.rda %>%
    scores(display='sites') %>%
    as.data.frame() %>%
    bind_cols(tenuis)
data.species.scores <- data.rda %>%
    scores(display='species') %>%
    as.data.frame() %>%
    mutate(Species=rownames(.))

# re-order treatment levels in data.sites.scores so, ambient, mid, high
data.sites.scores$Treatment <- factor(data.sites.scores$Treatment, levels = c('Ambient', 'Mid', 'High'))

# Making the probability circle for the plot
circle.prob <- 0.68
r <- sqrt(qchisq(circle.prob, df = 2)) * prod(colMeans(data.rda$CA$u[,1:2]^2))^(1/4)
theta <- c(seq(-pi, pi, length = 50), seq(pi, -pi, length = 50))
circle <- data.frame(PC1 = r * cos(theta), PC2 = r * sin(theta))

data.species.scores.sub <- data.species.scores %>%
  mutate(Length=sqrt(PC1^2 + PC2^2)) %>%
  filter(Length>r)  #if the length of the arrows is greater than the radius of 1 SE, we can remove species names
hjust <- ifelse(data.species.scores.sub$PC1>0,1,0)
vjust <- ifelse(data.species.scores.sub$PC2>0,0,1) #These are indicating which quadrant arrows are in to nudge in the right direction

#Put on some nice axes titles
eig <- eigenvals(data.rda)
paste(names(eig[2]), sprintf('(%0.1f%% explained var.)', 100 * eig[2]/sum(eig)))

```


All code needed to generate figure
```{r}
# cleaning up figure

gPCA <- ggplot() +
    geom_segment(data=NULL, aes(y=-Inf,x=0,yend=Inf,xend=0), # makes a plot
                 linetype='dotted')+
    geom_segment(data=NULL, aes(y=0,x=-Inf,yend=0,xend=Inf),
                 linetype='dotted')+ # makes the dashed cross
    geom_point(data=data.sites.scores, aes(y=PC2,x=PC1, col = Treatment))+ # adds the data
  scale_color_manual(values = c("#440154FF", "#FDE725FF", "#21908CFF"))+
    geom_segment(data=data.species.scores, aes(y=0,x=0,yend=PC2,xend=PC1), # adds the vectors
                 arrow=arrow(length=unit(0.3,'lines')), color='black')+
  geom_path(data = circle, aes(y=PC2,x=PC1),color = 'black', size = 1, alpha = 1)+ # adds probability circle
  geom_text(data=data.species.scores.sub, aes(y=PC2,x=PC1, label=Species),
                   hjust=hjust, vjust=vjust, color='black')+ # adds labels to vectors
  scale_y_continuous(paste(names(eig[2]), sprintf('(%0.1f%% explained var.)', 100 * eig[2]/sum(eig))))+
    scale_x_continuous(paste(names(eig[1]), sprintf('(%0.1f%% explained var.)', 100 * eig[1]/sum(eig))))+ # add nice axis titles
  theme_classic()+
  theme(axis.title.y =  element_text(family = "serif", size = 10),
        axis.title.x  =  element_text(family = "serif", size = 10), 
        axis.text.x = element_text(family = "serif", size = 10),
        axis.text.y = element_text(family = "serif", size = 10),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "right", 
        legend.text = element_text(family = "serif", size = 10), 
        legend.title = element_text(family = "serif", size = 10))
#ggsave(plot = g, file = "CleanPCA.jpeg", width = 15, height = 15, units = "cm")
```


# Correlation heatmap
```{r}
#load in some data 

data = tenuis %>% dplyr::select(all_of(physio))

# create a coorelation matrix of the data
 
# rounding to 2 decimal places
corr_mat <- round(cor(data),2) 
head(corr_mat)

# reorder corr matrix
# using corr coefficient as distance metric
dist <- as.dist((1-corr_mat)/2)
 
# hierarchical clustering the dist matrix
hc <- hclust(dist)
corr_mat <-corr_mat[hc$order, hc$order]

# reduce the size of correlation matrix
melted_corr_mat <- melt(corr_mat)
head(melted_corr_mat) # transforms data to 2 columns plus a coefficient

# plot
gcor <- ggplot(data = melted_corr_mat, aes(x=Var1, y=Var2,
                                   fill=value)) + 
  scale_fill_continuous(type = "viridis")+
  geom_tile()+
  geom_text(aes(Var2, Var1, label = value),
          color = "white", size = 3,  ) +
  labs(fill = "Cor coefficient")+
  theme(axis.title.y =  element_blank(), 
        axis.text.x = element_text(family = "serif", size = 10),
        axis.text.y = element_text(family = "serif", size = 10),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "right", 
        legend.text = element_text(family = "serif", size = 10),
        legend.title = element_text(family = "serif", size = 10))

#ggsave(plot = gcor, file = "CorPlot.jpeg", width = 20, height = 10, units = "cm")




#### Fixing the correlation heatmap to only include half the tiles

# Get lower triangle of the correlation matrix - helper functions
  get_lower_tri<-function(corr_mat){
    corr_mat[upper.tri(corr_mat)] <- NA
    return(corr_mat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(corr_mat){
    corr_mat[lower.tri(corr_mat)]<- NA
    return(corr_mat)
  }

# check that the helper functions worked, and that there's now NA's in one half of the matrix.  
upper_tri <- get_upper_tri(corr_mat)
upper_tri # yes, that worked. 


melted_cormat <- melt(upper_tri, na.rm = TRUE)

ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
# scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
 #  midpoint = 0, limit = c(-1,1), space = "Lab", 
  #  name="Pearson\nCorrelation") +
    scale_fill_continuous(type = "viridis", name = "Pearson\nCorrelation")+
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed() + 
  geom_text(aes(Var2, Var1, label = value), color = "white", size = 3) +
  theme(axis.title.y =  element_blank(), 
        axis.text.x = element_text(family = "serif", size = 10),
        axis.text.y = element_text(family = "serif", size = 10),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(fill = "transparent"),
        axis.line = element_blank(),
        legend.position = "right", 
        legend.text = element_text(family = "serif", size = 10),
        legend.title = element_text(family = "serif", size = 10))

#ggsave(plot = ggheatmap, file = "CorPlotNewHalf.jpeg", width = 20, height = 10, units = "cm")

```

```{r}
# combine PCA and corr heat map
g2 <- gPCA / ggheatmap
g2

ggsave(plot = g2, file = "PCACorHalfPlot.jpeg", width = 15, height = 18, units = "cm")
```

