---
title: "AcuteHeatMethodsPaper"
author: "JNielsen"
date: "28/01/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction 

This Rmarkdown produces the statistical outputs presented in the manuscript "Experimental considerations of acute heat stress assays to quantify coral thermal tolerance" by Nielsen et al 2022. This document follows the layout from the manuscript. 

##Libraries required for these analyses
```{r Libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(glmmTMB)    # for glmmTMB
library(sjPlot)     # for outputs and partial plots
library(car)        # for regression diagnostics
library(DHARMa)     # for residual diagnostics
library(effects)    # for partial effects plots
library(ggeffects)  # for effects plots in ggplot
library(multcomp)
library(dplyr)
library(MASS)
library(MuMIn)
library(rstatix)
library(emmeans)
library(lmtest)
library(multcompView)
library(patchwork)  # for arranging multiple plots
library(ggplot2)    # for graphing
library(ggfortify)  # for model diagnostics
library(gridExtra)
library(extrafont)  # for setting fonts in ggplot
loadfonts(device = "win") # needed to apply the fonts


# for RDA 
library(vegan)
library(GGally)
library(corrplot)
library(mvabund)
library(scales)
```

# Size analysis 

Here, we examine the effects of fragment size under heat stress. 
```{r}
# Load size data for both species, set factors
Size <- read.csv("Size.csv") %>%
  mutate(SampleID = as.factor(SampleID),
         Species = as.factor(Species),
         ClipID = as.factor(ClipID),
         RackID = as.factor(RackID),
         GenotypeID = as.factor(GenotypeID),
         Treatment = as.factor(Treatment),
         Tank = as.factor(Tank),
         Size = as.factor(Size))

Size$TreatSize <- paste(Size$Size, Size$Treatment) # for significant interaction terms later
Size$TreatSize = as.factor(Size$TreatSize)
summary(Size)
head(Size)
```

## Statistical analysis of fragment size effects in A. tenuis

Split the analysis into dealing with only one species at a time. 
```{r}
AtenSize <- Size %>% 
  filter(Species == "A. tenuis")
```


Chlorophyll content in A. tenuis
```{r}
# model
AtenChla <- glmmTMB(Chla_ug_cm2 ~ Treatment*Size + (1|GenotypeID),
                    data = AtenSize, 
                    family = gaussian(link = 'identity'), REML = TRUE)

# Diagnostics to determine if models satisfy assumptions
#Dharma residuals for glmm
chla.resid <- simulateResiduals(AtenChla, plot = TRUE)

summary(AtenChla)

# effects plots
plot_model(AtenChla, type = 'eff', terms =c('Treatment', 'Size'))
plot(allEffects(AtenChla), multiline=TRUE, ci.style='bars')

# R2 performance of model
performance::r2(AtenChla)

# summary figure
AtenChla.grid = with(AtenSize, list(Treatment=levels(Treatment),
                            Size=levels(Size)))
newdata = emmeans(AtenChla, ~Treatment*Size, at=AtenChla.grid,
                  offset=0, type='response') %>%
    as.data.frame
head(newdata)
ggplot(newdata, aes(y=emmean, x=Treatment)) +
  geom_pointrange(aes(ymin=lower.CL, ymax=upper.CL, color=Size),
                  position=position_dodge(width=0.2)) +
  scale_y_continuous('Chl-a content') +
  theme_bw()
```


### Cube-root transform function
```{r}
cbrt <- function(x) {
  sign(x) * abs(x)^(1/3)
}
```

Tissue colour change
```{r}
hist(AtenSize$ColourChange,breaks=100) # long left tail

AtenCol2 <- glmmTMB(cbrt(ColourChange) ~ Treatment*Size + (1|GenotypeID),
                    data = AtenSize, 
                    family = gaussian(link = 'identity'), REML = TRUE)
#Dharma residuals for glmm
col.resid2 <- simulateResiduals(AtenCol2, plot = TRUE)
testZeroInflation(col.resid2) # not zero-inflated
testDispersion(col.resid2) # looks fine, too.

summary(AtenCol2)

# effects plots
plot_model(AtenCol2, type = 'eff', terms =c('Treatment', 'Size'))
plot(allEffects(AtenCol2), multiline=TRUE, ci.style='bars')

# R2 performance of model
performance::r2(AtenCol2)

# post-hoc pairwise comparisons with significant interaction terms
emmeans(AtenCol2,  pairwise~Treatment|Size,  type='response',
        at=AtenSize$Size)

emmeans(AtenCol2,  pairwise~Size|Treatment,  type='response',
        at=AtenSize$Size) # need to run both ways around

# summary figure
AtenCol2.grid = with(AtenSize, list(Treatment=levels(Treatment),
                            Size=levels(Size)))
newdata = emmeans(AtenCol2, ~Treatment*Size, at=AtenCol2.grid,
                  offset=0, type='response') %>%
    as.data.frame
head(newdata)
ggplot(newdata, aes(y=response, x=Treatment)) +
  geom_pointrange(aes(ymin=lower.CL, ymax=upper.CL, color=Size),
                  position=position_dodge(width=0.2)) +
  scale_y_continuous('Colour change') +
  theme_bw()
```

Catalase activity 
```{r}
hist(AtenSize$CatalaseU,breaks=100) # large right tail 
# Model could be zero-inflated

# Fitting a gamma distribution with a log link due to nature of data

AtenCat2 <- glmmTMB(CatalaseU ~ Treatment*Size + (1|GenotypeID),
                    ziformula = ~Treatment, # specifying extra zeros in treatment. 
                    # zeros more often occur in high treatment
                    data = AtenSize, 
                    family = ziGamma(link = "log"), REML = TRUE)


#Dharma residuals for glmm
cat2.resid <- simulateResiduals(AtenCat2, plot = TRUE) # fixed it

summary(AtenCat2)

# R2 performance of model
performance::r2(AtenCat2)
```

Protein content
```{r}
hist(AtenSize$Protein_ug_cm2,breaks=100) # some right outliers

dat1 <- AtenSize %>% 
  group_by(Treatment) %>% 
  identify_outliers(Protein_ug_cm2) %>%
  filter(is.extreme == "TRUE") # 2 outliers identified

Size1 <- AtenSize %>% filter(!(SampleID %in% dat1$SampleID)) # outliers removed

hist(Size1$Protein_ug_cm2,breaks=100) # still long right tail and likely to be zero-inflated

# fit a zero-inflated model
AtenProt2 <- glmmTMB(Protein_ug_cm2 ~ Treatment*Size + (1|GenotypeID),
                    data = Size1,
                    ziformula = ~Treatment, 
                    family = ziGamma(link = 'log'), REML = TRUE)
#Dharma residuals for glmm
prot.resid2 <- simulateResiduals(AtenProt2, plot = TRUE) # yep, problem fixed.

summary(AtenProt2)

# R2 performance of model
performance::r2(AtenProt2)
```

## Prepare PAM data for analysis
Prepare PAM data for analysis. 
First have to calculate the Fv/Fm variable from the raw data F and Fm given in the spreadsheet. 
```{r}
PAM_Size <- read_csv("./data/PAM_SIZE.csv") %>%
  mutate(genus_species = paste0(substr(genus, start = 1, stop = 1), "_", species),
         tank = as.factor(tank),
         colonyID = as.factor(colonyID),
         replicate = as.factor(replicate),
         species = as.factor(species),
         size = as.factor(size),
         treatment = as.factor(treatment),
         collection.depth = as.numeric(collection.depth),
         Fm = ifelse(Fm <= F, F+1, Fm), # There should not be any Fm values < F
         FvFm = (Fm - F)/Fm)

# Subset to only A tenuis
PAM_AtenSize <- PAM_Size %>% 
  filter(species == "tenuis")
```

Fv/Fm stats for A. tenuis
```{r}
hist(PAM_AtenSize$FvFm,breaks=50) # slight left tail

dat1 <- PAM_AtenSize %>% 
  group_by(treatment) %>% 
  identify_outliers(FvFm) %>%
  filter(is.extreme == "TRUE") # 1 extreme outlier identified

Size1 <- PAM_AtenSize %>% filter(!(sampleID %in% dat1$sampleID)) # extreme outlier removed

hist(Size1$FvFm,breaks=50) # still slight left tail


AtenFvFm <- glmmTMB(FvFm ~ treatment*size + (1|colonyID),
                    data = Size1, 
                    family = gaussian(link = 'identity'), REML = TRUE)
#Dharma residuals for glmm
FvFm.resid <- simulateResiduals(AtenFvFm, plot = TRUE) 
# problem with Levene's test
testZeroInflation(FvFm.resid) # model NOT zero-inflated. 
testDispersion(FvFm.resid) # and dispersion looks fine.

# R2 performance of model
performance::r2(AtenFvFm)

summary(AtenFvFm)

```


# Size effect P. dam

```{r}
PdamSize <- Size %>%
  filter(Species == "P. damicornis")
summary(PdamSize)
```

Chlorophyll a content
```{r}
hist(PdamSize$Chla_ug_cm2,breaks=100) 

# model
PdamChla <- glmmTMB(Chla_ug_cm2 ~ Treatment*Size + (1|GenotypeID),
                    data = PdamSize, 
                    family = gaussian(link = 'identity'), REML = TRUE)

# Diagnostics to determine if models satisfy assumptions
#Dharma residuals for glmm
chla.resid <- simulateResiduals(PdamChla, plot = TRUE)

summary(PdamChla)

# effects plots
plot_model(PdamChla, type = 'eff', terms =c('Treatment', 'Size'))
plot(allEffects(PdamChla), multiline=TRUE, ci.style='bars')

# R2 performance of model
performance::r2(PdamChla) 

# post-hoc pairwise comparisons with significant interaction terms
emmeans(PdamChla,  pairwise~Treatment|Size,  type='response',
        at=PdamSize$Size)

emmeans(PdamChla,  pairwise~Size|Treatment,  type='response',
        at=PdamSize$Size) # need to run both ways around

# summary figure
PdamChla.grid = with(PdamSize, list(Treatment=levels(Treatment),
                            Size=levels(Size)))
newdata = emmeans(PdamChla, ~Treatment*Size, at=PdamChla.grid,
                  offset=0, type='response') %>%
    as.data.frame
head(newdata)
ggplot(newdata, aes(y=emmean, x=Treatment)) +
  geom_pointrange(aes(ymin=lower.CL, ymax=upper.CL, color=Size),
                  position=position_dodge(width=0.2)) +
  scale_y_continuous('Chl-a content') +
  theme_bw()
```
Tissue colour change
```{r}
hist(PdamSize$ColourChange,breaks=100) # long left tail, no obvious outliers

# model
# Transform on the response. 
PdamCol2 <- glmmTMB(cbrt(ColourChange) ~ Treatment*Size + (1|GenotypeID),
                    data = PdamSize, 
                    family = gaussian(link = 'identity'), REML = TRUE)

# Diagnostics to determine if models satisfy assumptions
#Dharma residuals for glmm
col.resid2 <- simulateResiduals(PdamCol2, plot = TRUE)

summary(PdamCol2)

# effects plots
plot_model(PdamCol2, type = 'eff', terms =c('Treatment', 'Size'))
plot(allEffects(PdamCol2), multiline=TRUE, ci.style='bars')

# R2 performance of model
performance::r2(PdamCol2) 

# post-hoc pairwise comparisons with significant interaction terms
emmeans(PdamCol2,  pairwise~Treatment|Size,  type='response',
        at=PdamSize$Size)

emmeans(PdamCol2,  pairwise~Size|Treatment,  type='response',
        at=PdamSize$Size) # need to run both ways around

# summary figure
PdamCol2.grid = with(PdamSize, list(Treatment=levels(Treatment),
                            Size=levels(Size)))
newdata = emmeans(PdamCol2, ~Treatment*Size, at=PdamCol2.grid,
                  offset=0, type='response') %>%
    as.data.frame
head(newdata)
ggplot(newdata, aes(y=response, x=Treatment)) +
  geom_pointrange(aes(ymin=lower.CL, ymax=upper.CL, color=Size),
                  position=position_dodge(width=0.2)) +
  scale_y_continuous('Colour change') +
  theme_bw()
```

Catalase 
```{r}
hist(PdamSize$CatalaseU,breaks=50) # huge outlier 

dat1 <- PdamSize %>% 
  group_by(Treatment) %>% 
  identify_outliers(CatalaseU) %>%
  filter(is.outlier == "TRUE") 
Size1 <- PdamSize %>% filter(!(SampleID %in% dat1$SampleID))
hist(Size1$CatalaseU,breaks=50) # could be zero inflated

# Fitting a different family due to nature of data; 
# no values can fall below zero. 
# so fitting a gamma distribution with a log link
# gamma-distribution model
PdamCat3 <- glmmTMB(CatalaseU ~ Treatment*Size + (1|GenotypeID),
                    data = Size1, 
                    ziformula = ~Treatment,
                    family = ziGamma(link = "log"), REML = TRUE)
cat.resid3 <- simulateResiduals(PdamCat3, plot = TRUE) # that's great
testZeroInflation(cat.resid3) # model is no-longer zero-inflated
testDispersion(cat.resid3) # and dispersion is good. 

summary(PdamCat3)

# R2 performance of model
performance::r2(PdamCat3) 

# post-hoc pairwise comparisons with significant interaction terms
emmeans(PdamCat3,  pairwise~Treatment|Size,  type='response',
        at=PdamSize$Size)

emmeans(PdamCat3,  pairwise~Size|Treatment,  type='response',
        at=PdamSize$Size) # need to run both ways around

# summary figure
PdamCat3.grid = with(PdamSize, list(Treatment=levels(Treatment),
                            Size=levels(Size)))
newdata = emmeans(PdamCat3, ~Treatment*Size, at=PdamCat3.grid,
                  offset=0, type='response') %>%
    as.data.frame
head(newdata)
ggplot(newdata, aes(y=response, x=Treatment)) +
  geom_pointrange(aes(ymin=lower.CL, ymax=upper.CL, color=Size),
                  position=position_dodge(width=0.2)) +
  scale_y_continuous('Catalase activity (U)') +
  theme_bw()
```
Protein content
```{r}
hist(PdamSize$Protein_ug_cm2,breaks=50) # likely to be zero-inflated; no obvious outliers

PdamProt <- glmmTMB(Protein_ug_cm2 ~ Treatment*Size + (1|GenotypeID),
                    data = PdamSize, 
                    family = gaussian(link = 'identity'), REML = TRUE)
#Dharma residuals for glmm
prot.resid <- simulateResiduals(PdamProt, plot = TRUE) 
# Levene's tests failed.
testZeroInflation(prot.resid) # model is zero-inflated
testDispersion(prot.resid) # but dispersion looks fine. 

# fit zero-inflated gamma
PdamProt2 <- glmmTMB(Protein_ug_cm2 ~ Treatment*Size + (1|GenotypeID),
                    ziformula = ~Treatment,
                    data = PdamSize, 
                    family = ziGamma(link = 'log'), REML = TRUE)
prot.resid2 <- simulateResiduals(PdamProt2, plot = TRUE)
testZeroInflation(prot.resid2) # Not zero-inflated 
testDispersion(prot.resid2) # dispersion is good. 

summary(PdamProt2)

# R2 performance of model
performance::r2(PdamProt2) 


# summary figure
PdamProt2.grid = with(PdamSize, list(Treatment=levels(Treatment),
                            Size=levels(Size)))
newdata = emmeans(PdamProt2, ~Treatment*Size, at=PdamProt2.grid,
                  offset=0, type='response') %>%
    as.data.frame
head(newdata)
ggplot(newdata, aes(y=response, x=Treatment)) +
  geom_pointrange(aes(ymin=lower.CL, ymax=upper.CL, color=Size),
                  position=position_dodge(width=0.2)) +
  scale_y_continuous('Protein content') +
  theme_bw()
```
Fv/Fm data
```{r}
# Subset to only P. damicornis
PAM_PdamSize <- PAM_Size %>% 
  filter(species == "damicornis")

hist(PAM_PdamSize$FvFm,breaks=50) #  right outlier; slight left tail

dat1 <- PAM_PdamSize %>% 
  group_by(treatment) %>% 
  identify_outliers(FvFm) %>%
  filter(is.outlier == "TRUE") # 3 outliers identified; all removed.

Size1 <- PAM_PdamSize %>% filter(!(sampleID %in% dat1$sampleID))


PdamFvFm <- glmmTMB(FvFm ~ treatment*size + (1|colonyID),
                    data = Size1, 
                    family = gaussian(link = 'identity'), REML = TRUE)
#Dharma residuals for glmm
fvfm.resid <- simulateResiduals(PdamFvFm, plot = TRUE) 
# Levene's and KS tests failed.
testZeroInflation(fvfm.resid) # model is NOT zero-inflated
testDispersion(fvfm.resid) # and dispersion looks fine. 

# R2 performance of model
performance::r2(PdamFvFm) 


summary(PdamFvFm)
```



# Time effect in percent change relative to control
## Strategy
Fit all time percent data with a generalised linear mixed effects model, assuming a Gaussian distribution. This will be checked with Dharma residuals. We include colonyID as a random effect to soak up some of the variation at each time point. Time is modeled as categorical as insufficient time points to model continuous. Additionally, trends do not appear linear nor display a constant rate of change over time. 


```{r ReadTime}
Time <- read.csv("./data/TimeComplex.csv")

Time <- Time %>%
  filter(TimePoint != "Before") %>% # remove the "before" time point which was only used to calculate initial colour of each fragment and is not used in the analyses below. 
    mutate(Timepoint.num = ifelse(TimePoint == "T0", 0, # this part of the code changes timepoint from factor to numeric
                                ifelse(TimePoint == "T1", 2, 
                                       ifelse(TimePoint == "T2", 6,
                                              ifelse(TimePoint == "T3", 10,
                                                     ifelse(TimePoint == "T4", 14,
                                                            ifelse(TimePoint == "T5", 24, 48))))))) %>%
   mutate(TimePoint = factor(TimePoint, levels =c("T0", "T1", "T2", "T3", "T4", "T5", "T6")))
summary(Time)

```

# Time stats with generalised linear mixed models

## Protein change
```{r}
TimeProt <- Time %>% 
  dplyr::select(ID, Genotype, TimePoint, Timepoint.num, Prot_percent_change) # subset to only one variable

# Outlier check
Time1 <- TimeProt %>% 
  group_by(TimePoint) %>% 
  identify_outliers(Prot_percent_change) %>%
  filter(is.outlier == "TRUE") #  4 outliers identified
  
Time2 <- TimeProt %>% filter(!(ID %in% Time1$ID)) # outliers removed

modelProt = glmmTMB(Prot_percent_change ~ TimePoint + (1|Genotype), data = Time2) # gaussian family unless specified

ModelProt.resid = simulateResiduals(modelProt, plot = TRUE) # something there isn't great

# so test for zero inflation
testZeroInflation(ModelProt.resid) # p value =1, not zero inflated.
# check for dispersion
testDispersion(ModelProt.resid) # not over/underdispersed either. 
# So issues likely due to complete lack of protein in last 2 timepoints. 

# partial plot
ggemmeans(modelProt, ~TimePoint) %>% plot(add.data=TRUE) 
newdata

# R2 performance of model
performance::r2(modelProt) # 0.48

# Model outputs
summary(modelProt)

# post-hoc comparisons with emmeans 
emmeans(modelProt, ~ TimePoint)%>%
  pairs()

# graph
newdata = emmeans(modelProt, ~ TimePoint, type = "response") %>% as.data.frame() %>%
 mutate(Timepoint.num = ifelse(TimePoint == "T0", 0, # this part of the code changes timepoint from factor to numeric
                                ifelse(TimePoint == "T1", 2, 
                                       ifelse(TimePoint == "T2", 6,
                                              ifelse(TimePoint == "T3", 10,
                                                     ifelse(TimePoint == "T4", 14,
                                                            ifelse(TimePoint == "T5", 24, 48))))))) 
newdata
ggplot(newdata, aes( y = emmean, x = Timepoint.num))+ 
  geom_pointrange(aes(ymin = lower.CL, ymax = upper.CL))+
  geom_point(data = Time2, aes(y=Prot_percent_change))


# save the emmeans in a spreadsheet 
write.csv(newdata, "./Outputs/ProtEmmeans.csv")
```

### Chlorophyll change 
```{r}
TimeChla <- Time %>% 
  dplyr::select(ID, Genotype, TimePoint, Timepoint.num, Chla_percent_change) 

# Outlier check
Time3 <- TimeChla %>% 
  group_by(TimePoint) %>% 
  identify_outliers(Chla_percent_change) %>%
  filter(is.outlier == "TRUE") #  2 outliers identified
  
Time4 <- TimeChla %>% filter(!(ID %in% Time3$ID)) 

modelChla = glmmTMB(Chla_percent_change ~ TimePoint + (1|Genotype), data = Time4)

ModelChla.resid = simulateResiduals(modelChla, plot = TRUE) 

# R2 performance of model
performance::r2(modelChla) # 0.34

# partial plot
ggemmeans(modelChla, ~TimePoint) %>% plot(add.data=TRUE) 

# Model outputs
summary(modelChla)

# post-hoc comparisons with emmeans 
emmeans(modelChla, ~ TimePoint)%>%
  pairs()

# graph
newdata = emmeans(modelChla, ~ TimePoint, type = "response") %>% as.data.frame() %>%
 mutate(Timepoint.num = ifelse(TimePoint == "T0", 0, # this part of the code changes timepoint from factor to numeric
                                ifelse(TimePoint == "T1", 2, 
                                       ifelse(TimePoint == "T2", 6,
                                              ifelse(TimePoint == "T3", 10,
                                                     ifelse(TimePoint == "T4", 14,
                                                            ifelse(TimePoint == "T5", 24, 48))))))) 
newdata
ggplot(newdata, aes( y = emmean, x = Timepoint.num))+ 
  geom_pointrange(aes(ymin = lower.CL, ymax = upper.CL))+
  geom_point(data = Time4, aes(y=Chla_percent_change))

# save the emmeans in a spreadsheet 
write.csv(newdata, "./Outputs/ChlaEmmeans.csv")
```

### Colour change
```{r}
TimeCol <- Time %>% 
  dplyr::select(ID, Genotype, TimePoint, Timepoint.num, Colour_percent_change) # subset to only one variable

# Outlier check
Time5 <- TimeCol %>% 
  group_by(TimePoint) %>% 
  identify_outliers(Colour_percent_change) %>%
  filter(is.outlier == "TRUE") #  1 outliers identified
  
Time6 <- TimeCol %>% filter(!(ID %in% Time5$ID)) 

modelCol = glmmTMB(Colour_percent_change ~ TimePoint + (1|Genotype), data = Time6) 

ModelCol.resid = simulateResiduals(modelCol, plot = TRUE) 

# R2 performance of model
performance::r2(modelCol) # 0.858

# partial plot
ggemmeans(modelCol, ~TimePoint) %>% plot(add.data=TRUE) 

# Model outputs
summary(modelCol)

# post-hoc comparisons with emmeans 
emmeans(modelCol, ~ TimePoint)%>%
  pairs()

# graph
newdata = emmeans(modelCol, ~ TimePoint, type = "response") %>% as.data.frame() %>%
 mutate(Timepoint.num = ifelse(TimePoint == "T0", 0, # this part of the code changes timepoint from factor to numeric
                                ifelse(TimePoint == "T1", 2, 
                                       ifelse(TimePoint == "T2", 6,
                                              ifelse(TimePoint == "T3", 10,
                                                     ifelse(TimePoint == "T4", 14,
                                                            ifelse(TimePoint == "T5", 24, 48))))))) 
newdata
ggplot(newdata, aes( y = emmean, x = Timepoint.num))+ 
  geom_pointrange(aes(ymin = lower.CL, ymax = upper.CL))+
  geom_point(data = Time6, aes(y=Colour_percent_change))

# save the emmeans in a spreadsheet 
write.csv(newdata, "./Outputs/ColEmmeans.csv")
```

### Fv/Fm
Just need to load a different spreadsheet here.
```{r}
Time_PAM_percent <- read.csv("./data//Time_PAM_Fv_percent.csv")
                             

TimeFvFm <- Time_PAM_percent %>% 
  dplyr::select(ID, colonyID, time, percent) 

# Outlier check
Time7 <- TimeFvFm %>% 
  group_by(time) %>% 
  identify_outliers(percent) %>%
  filter(is.outlier == "TRUE") #  3 outliers identified
  
Time8 <- TimeFvFm %>% filter(!(ID %in% Time7$ID)) %>% # outliers removed
  mutate(Timepoint.num = ifelse(time == "T0", 0, # this part of the code changes timepoint from factor to numeric
                                ifelse(time == "T1", 2, 
                                       ifelse(time == "T2", 6,
                                              ifelse(time == "T3", 10,
                                                     ifelse(time == "T4", 14,
                                                            ifelse(time == "T5", 24, 48))))))) %>%
  mutate(time = factor(time, levels =c("T0", "T1", "T2", "T3", "T4", "T5", "T6")))

# Model
modelFvFm = glmmTMB(percent ~ time + (1|colonyID), data = Time8) # gaussian family unless specified    

ModelFvFm.resid = simulateResiduals(modelFvFm, plot = TRUE) 

# R2 performance of model
performance::r2(modelFvFm) # 0.789

# partial plot
ggemmeans(modelFvFm, ~time) %>% plot(add.data=TRUE) 

# Model outputs
summary(modelFvFm)

# post-hoc comparisons with emmeans 
emmeans(modelFvFm, ~ time)%>%
  pairs()

# graph
newdata = emmeans(modelFvFm, ~ time, type = "response") %>% as.data.frame() %>%
 mutate(Timepoint.num = ifelse(time == "T0", 0, # this part of the code changes timepoint from factor to numeric
                                ifelse(time == "T1", 2, 
                                       ifelse(time == "T2", 6,
                                              ifelse(time == "T3", 10,
                                                     ifelse(time == "T4", 14,
                                                            ifelse(time == "T5", 24, 48))))))) 
newdata
ggplot(newdata, aes( y = emmean, x = Timepoint.num))+ 
  geom_pointrange(aes(ymin = lower.CL, ymax = upper.CL))+
  geom_point(data = Time8, aes(y=percent))

# save the emmeans in a spreadsheet 
write.csv(newdata, "./Outputs/FvFmEmmeans.csv")

```
### Catalase
```{r}
TimeCat <- Time %>% 
  dplyr::select(ID, Genotype, TimePoint, Timepoint.num, Catalase_percent_change) # subset to only one variable

# Outlier check
Time9 <- TimeCat %>% 
  group_by(TimePoint) %>% 
  identify_outliers(Catalase_percent_change) %>%
  filter(is.outlier == "TRUE") #  6 outliers identified
  
Time10 <- TimeCat %>% filter(!(ID %in% Time9$ID)) 

modelCat = glmmTMB(Catalase_percent_change ~ TimePoint + (1|Genotype), data = Time10) 

ModelCat.resid = simulateResiduals(modelCat, plot = TRUE)  # not great, and I need to try and do something about those
# so test for zero inflation
testZeroInflation(ModelCat.resid) # p value =1, not zero inflated.
# check for dispersion
testDispersion(ModelCat.resid) # not over/underdispersed either. 


# R2 performance of model
performance::r2(modelCat) # 0.596

# partial plot
ggemmeans(modelCat, ~TimePoint) %>% plot(add.data=TRUE) 

# Model outputs
summary(modelCat)

# post-hoc comparisons with emmeans 
emmeans(modelCat, ~ TimePoint)%>%
  pairs()

# graph
newdata = emmeans(modelCat, ~ TimePoint, type = "response") %>% as.data.frame() %>%
 mutate(Timepoint.num = ifelse(TimePoint == "T0", 0, # this part of the code changes timepoint from factor to numeric
                                ifelse(TimePoint == "T1", 2, 
                                       ifelse(TimePoint == "T2", 6,
                                              ifelse(TimePoint == "T3", 10,
                                                     ifelse(TimePoint == "T4", 14,
                                                            ifelse(TimePoint == "T5", 24, 48))))))) 
newdata
ggplot(newdata, aes( y = emmean, x = Timepoint.num))+ 
  geom_pointrange(aes(ymin = lower.CL, ymax = upper.CL))+
  geom_point(data = Time10, aes(y=Catalase_percent_change))

# save the emmeans in a spreadsheet 
write.csv(newdata, "./Outputs/CatEmmeans.csv")

```


# Assay selection - PCA preparation
```{r}
#Import Data
tenuis = read.csv("./data/TenuisMDS.csv") %>% 
  mutate(Reef = as.factor(Reef),
         Species = as.factor(Species),
         Treatment = as.factor(Treatment))
tenuis 
```
Preparing data for PCA (which turns in to Redundancy Analysis)
```{r}
#outliers
tenuis %>% filter(Reef == "Corbett") %>% arrange(desc(CatalaseU)) # very large catalase outlier in sample 891
tenuis <- tenuis %>% 
  filter(SampleID != "891")
```

Isolate traits in a separate data frame
```{r}
physio = c("Chla_ug_cm2", "Colour.Change", "CatalaseU", "Protein_ug_cm2", "FvFm")
```

Examining the correlations between traits. 
The larger and more blue dots indicate higher correlation between variables. 
Here, Colour change and FvFm are highly correlated while colour change is also correlated to chl a content and protein content. FvFm is also correlated with chl a and protein content. 
```{r}
tenuis %>% dplyr::select(all_of(physio)) %>% 
  cor %>% 
  corrplot(type = "upper", diag = F)
```

```{r message = FALSE}
ggpairs(tenuis %>% dplyr::select(all_of(physio)), lower = list(continuous = "smooth"), 
        diag = list(continuous = "density"),
        axisLabels = "show")
```


##Redundancy analysis (PCA)
This is where the correlations between variables above matter. Because these variables are correlated, we need to do a redudancy analysis rather than a PCA. 
```{r}
data.rda <- rda(tenuis %>% dplyr::select(all_of(physio)), scale=TRUE)  # based on correlation, pca is an unconstrained redundancy analysis. 
summary(data.rda, display=NULL)
```
With correlation, inertia will simply be the number of variables (species)
Of the 6 units of variance (inertia), the first (PC1) explains 43% of the original variability in the data
Collectively PC1 and PC2, explain 63% of the variance. 

```{r}
screeplot(data.rda)
abline(a=1,b=0)
```
Here 5 variables, so if they were all uncorrelated they would explain 1/6th of the data. In a scaled pca, if an eigenvalue is contributed more than 1 inertia, it is contributing more than its share. SO here you might keep 2. 

```{r}
scores(data.rda, choices=1:3, display='sites')
scores(data.rda, choices=1:3, display='species')
biplot(data.rda, scaling='species')
biplot(data.rda, scaling='sites')
```

```{r}
data.sites.scores <- data.rda %>%
    scores(display='sites') %>%
    as.data.frame() %>%
    bind_cols(tenuis)
data.species.scores <- data.rda %>%
    scores(display='species') %>%
    as.data.frame() %>%
    mutate(Species=rownames(.))

# re-order treatment levels in data.sites.scores so, ambient, mid, high
data.sites.scores$Treatment <- factor(data.sites.scores$Treatment, levels = c('Ambient', 'Mid', 'High'))

# Making the probability circle for the plot
circle.prob <- 0.68
r <- sqrt(qchisq(circle.prob, df = 2)) * prod(colMeans(data.rda$CA$u[,1:2]^2))^(1/4)
theta <- c(seq(-pi, pi, length = 50), seq(pi, -pi, length = 50))
circle <- data.frame(PC1 = r * cos(theta), PC2 = r * sin(theta))

data.species.scores.sub <- data.species.scores %>%
  mutate(Length=sqrt(PC1^2 + PC2^2)) %>%
  filter(Length>r)  #if the length of the arrows is greater than the radius of 1 SE, we can remove species names
hjust <- ifelse(data.species.scores.sub$PC1>0,1,0)
vjust <- ifelse(data.species.scores.sub$PC2>0,0,1) #These are indicating which quadrant arrows are in to nudge in the right direction

```




#Correlations 
All correlations were generated using the ggheatmap function in the Figures R script. 